### 2018-03-11 paper/7-related.tex
Rename the chapters so they are in order
Write the related work

\begin{itemize}
    \item Memoization
    \item Self-adjusting computation
    \item Lens-like formulation
    \item Other build systems: \Ninja, \cmd{nix},
          cloud (Blaze -- first? \Buck, Pants, CloudBuild, XXX),
          Shake derivatives (Jane Street's \cmd{Jenga}, Apple's \cmd{swift-llbuild}).
\end{itemize}

John Ericson suggested in a blog comment that nix may be somewhat monadic, see:
\url{https://blogs.ncl.ac.uk/andreymokhov/cloud-and-dynamic-builds/\#comment-1849}.

>>>>

In section \S\ref{sec-background} we cover many build systems, but there is frustrating little work comparatively analysing build systems. In this section we cover a few build systems that may not

\subsection{Other Build Systems}

Most build systems, when viewed at the level we talk, can be captured with minor variations on the code presented in \S\ref{sec-implementations}. As some examples:

\begin{itemize}
\item \Ninja \cite{ninja} combines the dependency strategy of \Make with the validating traces of \Shake - our associated implementation provides such a combination. Ninja is also capable of modelling rules that produce multiple results, a limited form of polymorphism \S\ref{sec-polymorphism}.
\item \Tup \cite{tup} functions much like \Make, but with a refined dirty-bit implementation that watches the file system for changes and can thus avoid rechecking the entire graph, and with automatic deleting of stale results.
\item \Redo \cite{redo} almost exactly matches \Shake at the level of detail given here, differing only on aspects like polymorphic dependencies \S\ref{sec-polymorphism}.
\item \Buck \cite{buck} and \CloudBuild \cite{cloudbuild} are both fundamentally similar to \Bazel, differing around areas such as non-determinism \S\ref{sec-sec-non-determinism}.
\item \Nix \cite{nix} has coarse-grained monadic dependencies, with precise hashing of dependencies and downloading of precomputed build products, so may be, at some level of abstraction, the same as the hypothetical Cloud Shake from \S\ref{sec-implementation-cloud-shake}. However, \Nix is not intended as a build system, and the coarse grained nature (packages, not individual files) makes it targetted to a different purpose. \todo{NM}{Read comment below when not on an airplane.}
% John Ericson suggested in a blog comment that nix may be somewhat monadic, see:
% \url{https://blogs.ncl.ac.uk/andreymokhov/cloud-and-dynamic-builds/\#comment-1849}.
\item \Pluto \cite{pluto} has power to express cyclic dependenies, but that is implemented \todo{NM}{Read the paper so I can say something precise}
\end{itemize}

The one build system we are aware of that cannot be modelled in our framework is \Fabricate \cite{fabricate}. In \Fabricate a build system is a script which is run in-order, in the spirit of\footnote{\Fabricate requires scripts to be written in Python, but those details are not fundamental to what makes \Fabricate special.}:

\begin{minted}{bash}
gcc -c util.c
gcc -c main.c
gcc util.o main.o -o main.exe
\end{minted}

To achieve minimality, each separate command is traced at the OS-level, allowing \Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from \cmd{util.c}. In future runs \Fabricate runs the script from start to finish, skipping any commands where no inputs have changed.

Taking our abstraction, it is possible to encode \Fabricate assuming that nodes like \cmd{gcc -c util.c} are keys, there is a linear dependency between each successive node, and that the OS-level tracing can be lifted back as a monadic \hs{Compute} function\footnote{In fact, \Shake has an execution mode that can model \Fabricate{}-like build systems -- see \hs{Development.Shake.Forward} in the \Shake library.}. However, in our pure model the mapping is not perfect as \hs{gcc} writes to arbitrary files whose locations are not known in advance.

\subsection{Memoization}

\todo{AM}{Not convined this is really related, but if you think it is, please write something.}

\subsection{Self-adjusting computation}

While not typically considered build systems, self-adjusting computation is a well studied area, and in particular the constrast between different formulations has been well studied. \todo{AM}{Say something.}

\subsection{Profunctor Optics}

The \hs{Compute} definition of:

\begin{minted}{haskell}
type Compute c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}

Looks tanilisingly close to the profunctor optics definition in \cite{gibbons-profunctor-optics}:

\begin{minted}{haskell}
type Optic c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> f v
\end{minted}

The properties of such optics are well studied, and the functions like \hs{getDependencies} are very much based on observations from that field of work. Alas, we have been unable to remove the \hs{Maybe} used to encode whether a file is an input, without complicating other aspects of our definition. Furthermore, the \hs{Build} operation lacks any further such symmetry.
